---
title: "Coursera - Machine Learning"
subtitle: "Week 4 Peer Assignement"
author: "Toby Jones"
date: "September 4, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(caret)
library(randomForest)
```

# Data Loading
```{r}
# see preperation.R for the downloading of the files from their internet location
training <- read.csv("data/pml-training.csv",row.names=1,na.strings = "NA")
testing <- read.csv("data/pml-testing.csv",row.names=1,na.strings = "NA")
```

# Preprocessing
```{r}
# Remove near zero variance predictors
nzv <- nearZeroVar(training, saveMetrics = TRUE)
training <- training[,!nzv$nzv]
testing <- testing[,!nzv$nzv]

# Remove variables with missing values
training_filter_na <- training[,(colSums(is.na(training)) == 0)]
testing_filter_na <- testing[,(colSums(is.na(training)) == 0)]

# Remove unnecessary columns
colRm_train <- c("user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","num_window")
colRm_test <- c("user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","num_window","problem_id")
training_colRm <- training_filter_na[,!(names(training_filter_na) %in% colRm_train)]
testing_colRm <- testing_filter_na[,!(names(testing_filter_na) %in% colRm_test)]
dim(training_colRm)
dim(testing_colRm)
```

Split training data into training set and validation set
```{r}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
training_clean <- training_colRm[inTrain,]
validation_clean <- training_colRm[-inTrain,]
```

Test Linear Regression Model
```{r}
cor <- abs(sapply(colnames(training_clean[, -ncol(training)]), function(x) cor(as.numeric(training_clean[, x]), as.numeric(training_clean$classe), method = "spearman")))
```

## Random Forest Model

Fit the data within a random forest model and check the model performance against the validation set.

```{r}
set.seed(1234)
# Fit rf model
rfFit <- train(classe ~ ., method = "rf", data = training_clean, importance = T, trControl = trainControl(method = "cv", number = 4))
validation_pred <- predict(rfFit, newdata=validation_clean)
# Check model performance
confusionMatrix(validation_pred,validation_clean$classe)
```

```{r}
# Check important variable
imp <- varImp(rfFit)$importance
varImpPlot(rfFit$finalModel, sort = TRUE, type = 1, pch = 19, col = 1, cex = 1, main = "Importance of the Predictors")
```

This algorithm generates a model with 99.35% accuracy.  The top four predictors are 'yaw_belt', 'magnet_dumbell_z', 'roll_belt', and 'pitch_belt'.

# Prediction
The last step is the run the prediction on the test set and write out the outcomes.
```{r}
predictions  <- predict(rfFit, newdata=testing_colRm)
print(predictions)
```

# Generate Answer Files to submit for Assignment
```{r}
if(!file.exists("./answers")){dir.create("./answers")}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("./answers/problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictions)
```